Analysis of the Anomalous Doppler Effect from Quantum Theory to Classical Dynamics Simulations

### 2025-05-11 paper modification/CPB-tex-template/CPB Template-LaTeX/paper_v2.tex
after first review

The physics of ADE was originally explained using quantum analysis by Frank and Ginzburg \cite{frank1960optics,ginzburg1979nonlinear}. In this paper, we build on Ginzburg’s quantum framework by incorporating angular momentum conservation to provide a more detailed analysis of ADE. This approach offers new insights into the relationship between wave angular momentum and ADE. Despite the model's simplicity, to the best of our knowledge, the conservation of angular momentum in the context of ADE has not been previously analyzed or discussed.

Furthermore, we conduct numerical simulations of a single electron resonating with an electromagnetic (EM) wave in uniform static electric and magnetic fields, using classical dynamical equations. These simulations reveal the relationship between wave angular momentum and the resonance condition. The energy transfer ratio from the electron’s translational kinetic energy to its gyrokinetic energy during resonance is also computed, and the results are found to be in strong agreement with quantum theoretical predictions.

The remainder of this paper is organized as follows: Section~\ref{Quantum}~~~ presents the quantum analysis incorporating angular momentum conservation. Section III details the numerical setup and methodology, including the time evolution of velocity and kinetic energy. Section IV discusses the energy transfer ratio and polarization characteristics. Finally, Section V provides a brief discussion and conclusion.

@@

Let's consider the process in which an electron cyclotron system under a uniform magnetic field emits a photon, as shown in Fig.~3. The moving electron has the velocity $v_z$ along the background magnetic field and the $v_\perp$ cyclotron velocity. The kinetic energy along $z$ is $T = \gamma m_e c^2 - m_e c^2$, where $\gamma$ refers to the Lorentz factor. The internal energy represents as $U = \frac{1}{2} \gamma m_e v_\perp^2$.

@@

The $z$-component of the orbital angular momentum operator can be expressed in spherical coordinates as:
\begin{equation}
\hat{L}_z = -i\hbar\frac{\partial}{\partial\phi} \Psi \label{eq:Lz}
\end{equation}
Combining Eq.~(\ref{eq:psi}) with Eq.~(\ref{eq:Lz}), we have 
\begin{equation}
-i\hbar \frac{\partial}{\partial \phi} \Psi = (p_\phi - eA_\phi) r \Psi
\end{equation}
As a result, the eigenvalue of  $L_z $can be expressed as
\begin{equation}
L_z = (p_\phi - e A_\phi) r  \label{eq:Lz_2}
\end{equation}
With \( p_\phi = \gamma m_e v_\perp \), \( A_\phi = \dfrac{r B_0}{2} \), and \( r = \dfrac{\gamma m_0 v_\perp}{B_0 e} \), Eq.~(\ref{eq:Lz_2}) can be rewritten as:
\begin{equation}
L_z = \frac{1}{2} \cdot \frac{\gamma m_0 v_\perp^2}{\omega_{ce}} = \frac{U}{\omega_{ce}},
\end{equation}
where \(\omega_{ce} = \dfrac{eB}{m_0 \gamma} = \dfrac{\omega_0}{\gamma}\) and \(U = \dfrac{1}{2} \gamma m_0 v_\perp^2\).
Here, \( m_0 \) is the electron rest mass, \( \gamma \) is the Lorentz factor, and \( \omega_0 \) is the electron cyclotron frequency in the rest frame (\( \omega_0 > 0 \)). The conservation of angular momentum in the \( z \)-direction is expressed as
\[
L_{z2} + m\hbar = L_{z1}.
\]
The variation in the angular momentum of the electron along the \( z \)-axis is given by:
\begin{equation}
\Delta L_{21} = L_{z2} - L_{z1} = \frac{U_2 - U_1}{\omega_{ce}} = -m\hbar \label{eq:DeltaL21}
\end{equation}
Here, \( m \) is the quantum number of the photon's angular momentum in the \( z \)-direction. The internal energy change is given by \( \Delta U_{21} = U_2 - U_1 \). With Eq.~(\ref{eq:DeltaL21}), it can be transformed as:
\begin{equation}
\Delta U_{21}=-m \hbar \omega_{ce}  \label{eq:DeltaU21}
\end{equation}
According to the Eq.~(\ref{eq:DeltaU00})  and Eq.~(\ref{eq:DeltaU21}), the change in electron energy  could be presented as 
\begin{equation}
\hbar \vec{k} \cdot \vec{v} = \hbar \omega - m \hbar \omega_{ce}
\end{equation}
This result is consistent with previous findings~\cite{tamm1959general, frank1960optics, nezlin1976negative, coppi1976slide, frolov1986excitation, ginzburg1996radiation}. Here, \( \hbar \vec{k} \cdot \vec{v} \) represents the loss of kinetic energy \( \Delta T_{12} \), \( \hbar \omega \) represents the energy of the photon, and \( -m \hbar \omega_{ce} \) represents the change in the electron gyrokinetic energy \( \Delta U_{21} \) (the internal energy change). The ratio between the internal energy change \( \Delta U_{21} \) and the kinetic energy change \( \Delta T_{12} \) can be expressed as
\begin{equation}
\frac{\Delta U_{21}}{\Delta T_{21}} = \frac{m \hbar \omega_{ce}}{\hbar \vec{k} \cdot \vec{v}}  \label{eq:DU21DT21}
\end{equation}
 This results is a critical criterion to compare with the classical dynamic simulation in the section 2. It is also proved based on classical theory in the Appendix.
After simpifying the Eq.~(\ref{eq:DU21DT21}), we finally have the classical wave-particle resonant condition 
\begin{equation}
\omega = k_z v_z + m \omega_{ce}
\end{equation}
The variable \( m \) represents the quantum number associated with the angular momentum of the photon. Since a photon possesses both orbital angular momentum (\( l\hbar\), where \(l = 0,\pm1,\pm2,\pm3,... \)) and intrinsic spin angular momentum (\( s\hbar \), where \( s = \pm 1 \))~\cite{arnaut2000orbital}, the total angular momentum can be expressed as \( m\hbar = l\hbar + s\hbar \).

If we consider a plane wave, only the spin angular momentum contributes in this context (i.e., \( l = 0 \)). Therefore, there are two possible scenarios regarding the sign of \( m \), corresponding to the two possible spin states of the photon: \( m = +1 \) or \( m = -1 \).

\begin{enumerate}
\item  	For $m>0$, where $\Delta U_{21}<0$, the internal energy of the cyclotron electron decreases after emitting a photon. If the angular momentum quantum number $m =1$, the emitted photon exhibits right-hand circular polarization. This process is known as the NDE.
\item    	For $m<0, \Delta U_{21}>0$, the cyclotron electron gains internal energy after emitting a photon. The emission photo will have left-hand circular polarization if the angular momentum quantum number $m = -1$\footnote{The difference between the definition of left- and right-hand polarization for \( m \) and the conventional definition~\cite{kiang2008angular} arises from the choice of \( \omega_0 > 0 \) used here, where \( m > 0 \) corresponds to the same sense of rotation as the electron's right-hand motion, which corresponds to right-hand polarization for \( \vec{k} \parallel \vec{B}_0 \).
}. This process is known as the ADE. 
\end{enumerate}

While ADE and NDE describe spontaneous emission phenomena that occur without external field intervention, in our simulation model an external electromagnetic (E.M) waves is introduced as resonant fields interacting with electrons in static magnetic and electric fields. This approach provides a framework for analysing ADE under resonant conditions, referred to here as Anomalous Doppler Resonance (ADR). Under such resonance, both emission and absorption processes can occur, depending on the phase relationship between the electron’s perpendicular velocity and the electric field of the E.M wave. A detailed analysis is provided in the Appendix.

Depsite nonlinear analyses of electron interactions with E.M waves—excluding static electric fields—have been presented in numerous studies \cite{liu2004particle,qian1999exact,weyssow1990motion,gogoberidze2005origin,roberts1964motion,bourdier2000dynamics,nusinovich1999theory,nusinovich1995theory,qian2000relativistic}, fewer investigations have considered the influence of a static electric field during resonance with E.M waves. Due to the complexity of the nonlinear processes involved, analytical solutions are nearly impossible to obtain, making numerical simulations essential in this context.

\section{Classical dynamic simulation of ADR}\label{sec:level2}
The ADE process has been analyzed based on quantum theory, demonstrating that the angular momentum of the emitting photon determines the resonance condition. Specifically, only angular momentum with  $m < 0 $ corresponds to the ADE process, while $m >0 $ corresponds to the NDE process. These characteristics will be tested through the interaction of E.M wave and the electron during ADR and Normal Doppler Resonance (NDR), and the energy transfer ratio can also be verified through numerical simulations.
\subsection{Numerical simulation setup}
To analyze the resonant process from the perspective of classical dynamics and to provide a direct comparison between quantum and classical dynamic results, the following scenario is considered: A uniform magnetic field \( \vec{B}_0 \) is applied along the \( z \)-direction. An electrostatic field \( \vec{E}_0 \), oriented in the opposite direction to \( \vec{B}_0 \) (as illustrated in Fig.~4), is used to accelerate the electron. 
\begin{figure}
\centering
\includegraphics[width=0.4\textwidth]{Figure4.eps}% Here is how to import EPS art
\caption{\label{fig:Setup}The uniform static magnetic field is set along the z axis, the electrostatic field $E_0$ is oriented opposite to the $B_0$ field, and the wavevector $k$ is aligned parallel to the $B_0$ field.}
\end{figure}
We consider the interaction between an electron entering the system with velocity \( v_z \), parallel to the magnetic field \( B_0 = B_z \), and a linearly or circularly polarized transverse electromagnetic (TEM) wave propagating in a homogeneous dielectric medium with a refractive index \( n > 1 \).

The induced linearly polarized wave along \( \vec{B}_0 \) can be decomposed into a combination of a right-hand circularly polarized wave (\( m = 1 \)) and a left-hand circularly polarized wave (\( m = -1 \)), such that
\(
\vec{E}_w = \vec{E}_R + \vec{E}_L,
\)
where
\(
\vec{E}_R = \frac{1}{2} E_0 (\vec{e}_x + i\vec{e}_y) \exp[i(\vec{k} \cdot \vec{r} - \omega t)],
\vec{E}_L = \frac{1}{2} E_0 (\vec{e}_x - i\vec{e}_y) \exp[i(\vec{k} \cdot \vec{r} - \omega t)].
\)
If the wavevector \( \vec{k} \) lies in the \( y \)-\( z \) plane with a crossing angle \( \theta_k \) relative to the \( z \)-axis, then the new coordinate unit vectors for the wave field expression should be rotated accordingly to align with the direction of \( \vec{k} \). The transformed basis vectors are:
\begin{equation}
\begin{pmatrix}
\vec{e}_x' \\
\vec{e}_y'
\end{pmatrix}
=
\begin{pmatrix}
1 & 0 & 0 \\
0 & \cos\theta_k & \sin\theta_k
\end{pmatrix}
\cdot
\begin{pmatrix}
\vec{e}_x \\
\vec{e}_y \\
\vec{e}_z
\end{pmatrix}
\end{equation}
The magnetic field of E.M wave is
\begin{equation}
\vec{B}_w = \frac{\vec{k} \times \vec{E}_w}{\omega}
\end{equation}

@@

To simulate the evolution of $\bm{r}$ and $\bm{p}$, the above system is discretized using the Volume-Preserving Algorithm. Let $k$ denote the iteration step and $\text{Cay}(\bm{A})$ represent the Cayley transform of matrix $\bm{A}$:

\begin{equation}
\left\{
\begin{aligned}
\bm{r}^*_{k+\frac{1}{2}} &= \bm{r}^*_k + \frac{\Delta t^*}{2 \gamma_k} \bm{p}^*_k, \\
\bm{p}^{*-} &= \bm{p}^*_k + \frac{\Delta t^*}{2} \bm{E}^*_{k+\frac{1}{2}}, \\
\bm{p}^{*+} &= \text{Cay} \left( \frac{\Delta t^* \hat{\bm{B}}^*}{2 \gamma^{*-}} \right) \bm{p}^{*-}, \\
\bm{p}^*_{k+1} &= \bm{p}^{*+} + \frac{\Delta t^*}{2} \bm{E}^*_{k+\frac{1}{2}}, \\
\bm{r}^*_{k+1} &= \bm{r}^*_{k+\frac{1}{2}} + \frac{\Delta t^*}{2 \gamma_{k+1}} \bm{p}^*_{k+1},
\end{aligned}
\right.
\end{equation}
The dimensionless parameters are momentum \( p^* = p/(m_0 c) \), magnetic field \( B^* = B/(e \tau_{ce} m_0) \), total electric field \( E^* = E/(\frac{m_0 c}{\tau_{ce} e}) \), time step \( \Delta t^* = \Delta t / \tau_{ce} \), and position \( r^* = r / (\tau_{ce} c) \) respectively, where the \( \tau_{ce} \) is the electron cyclotron period (\( \tau_{ce} = 2\pi / \omega_{ce} \)) and \( \gamma^* = \sqrt{1 + p^{*2}} \) is Lorentz factor. The dimensionless magnetic matrix \( \mathbf{B}^* \) \cite{zhang2015volume} is written as

@@

The evolution of the electron’s motion is shown in fig. 5. As the electron accelerates from stationary in the electrostatic field (fig. 5(b)), the resonant frequencies increase simultaneously (Fig.~5(a)). The change of parallel velocity caused by electromagnetic wave can be quantified as \( \Delta v = v_z - v_{zE0} \) as shown in Fig.~5(c), where \( v_z \) represents the parallel velocity under the given scenario, while \( v_{zE0} \) denotes the parallel velocity resulting solely from the electrostatic field, which can be calculated using a theoretical equation as

@@

The cyclotron velocity is shown in fig. 5(d). The work done by electromagnetic wave is shown in fig. 5(e), which can be calculated by integrating the power with time as 
\(
E_{\parallel \text{emw}} = \int P_{\parallel \text{emw}} dt ,
\)
and 
\(
P_{\parallel \text{emw}} = -e \vec{v}_\perp \times \vec{B}_{\perp \text{emw}} \cdot v_z.
\)
Since all discrete data points are available from the simulation, it is not difficult to integrate all the discrete data over time. Fig. 5(f) shows the gyro-kinetic energy evolution with time, where 
\(
E_\perp = \frac{1}{2} m_e v_\perp^2.
\)

\begin{figure}[ht]
\centering
\includegraphics[width=0.8\textwidth]{Figure5.pdf}% Here is how to import EPS art
\caption{\label{fig:NumSim}Kinetic evolution of electrons in a magnetic field with an electromagnetic wave during acceleration. 
    (a) Frequencies of ADE, NDE, and source wave frequency. $\theta_k$ refers to the angle between $\vec{k}$ and the $z$ axis, here $\theta_k = 0$. 
    (b) The parallel velocity $v_z$ in the case with and without the electromagnetic wave. 
    (c) The change of parallel velocity caused by the electromagnetic wave. 
    (d) The cyclotron velocity $v_\perp$. 
    (e) The parallel kinetic energy transferred to the electron by the electromagnetic wave. 
    (f) The evolution of gyro-kinetic energy.}
    \label{fig:kinetic_evolution}
\end{figure}

@@

As shown in Fig.~5(a), around \( 23\tau_{ce} \), the normal Doppler frequency matches that of the induced wave, leading to a rapid increase in the cyclotron velocity \( v_{\perp} \) (Fig.~5(b)). Simultaneously, the change in parallel velocity induced by the electromagnetic wave also increases. This phenomenon can be interpreted as the electron cyclotron system absorbing a photon during the Normal Doppler Effect, resulting in an increase in both parallel kinetic energy and cyclotron energy. The change in parallel kinetic energy caused by the electromagnetic wave is shown in Fig.~5(e), where \( \Delta T_{21} = \Delta E_{||\text{emw}} \approx 5.4 \times 10^{-5} \, \text{eV} \). The increase in cyclotron energy is shown in Fig.~5(e), where \( \Delta U_{21} = \Delta E_{\perp} \approx 10.7 \times 10^{-5} \, \text{eV} \). The energy transfer ratio between internal energy and kinetic energy during resonance is given by \( \frac{\Delta U_{21}}{\Delta T_{21}} \approx 1.98 \). According to quantum theory, the energy ratio is given by Eq.~(13). Here \( m = 1 \) for NDE and \( k = 10^5 \, \text{m}^{-1} \) along the \( z \)-axis, the resonant velocity \( v_z \approx 19 \times 10^3 \, \text{m/s} \) and \( \omega_{ce} \approx 3.51 \times 10^9 \, \text{s}^{-1} \). Finally, \( n_p = 1.85 \), which is in close agreement with the simulation results.

>>>>

The physics of the Anomalous Doppler Effect (ADE) was first elucidated through quantum analysis by Frank and Ginzburg \cite{frank1960optics,ginzburg1979nonlinear}. In this work, we extend Ginzburg's quantum framework by incorporating  the conservation of angular momentum to provide a more rigorous analysis of ADE. This approach yields new insights into the relationship between wave angular momentum and ADE under resonant conditions, which it is referred to Anomalous Doppler Resonance (ADR). For an electron moving in a magnetic field and interacting with an external electromagnetic (EM) wave, the general resonance condition is given by:
\(
    \omega = m\omega_{ce} + \vec{k} \cdot \vec{v},
\)
where $\vec{k}$ is the wave vector, $\omega_{ce}$ is the electron cyclotron frequency (here $\omega_{ce}>0$ ), $\vec{v}$ is the electron velocity, and $\omega$ is the wave angular frequency, $m = 0, \pm1, \pm2, \pm3, \ldots$ represent the Landau level\cite{coppi1976slide}. Specially,  for plane EM waves, we find that resonance is restricted to the fundamental harmonics ($m = \pm1$) due to spin angular momentum conservation, reducing the condition to
\(
    \omega = \pm\omega_{ce} + \vec{k} \cdot \vec{v}.
\)
where the negative sign refers to ADR condition while the positive sign refers to Normal Doppler Resonance (NDR) condition. This represents a significant constraint compared to previous theoretical treatments\cite{dendy1987classical} which suggested possible resonance at all harmonic orders ($m = \pm1, \pm2, \dots$).
Despite the  simplicity of the  model, our analysis demonstrates that angular momentum conservation plays a crucial role in EM wave-electron resonance - an aspect that, to the best of our knowledge, has not been previously addressed in the literature.

Furthermore, we perform numerical simulations of a single electron interacting resonantly with an EM wave in the presence of uniform static electric and magnetic fields, using classical equations of motion.These simulations elucidate the relationship between wave angular momentum and the resonance mechanism. Additionally, we compute the energy transfer ratio from the electron’s translational kinetic energy to its gyrokinetic energy during resonance, and the results show strong agreement with predictions from quantum theory.

The remainder of this paper is organized as follows. Section~\ref{Quantum}~~~ develops the quantum theoretical framework incorporating angular momentum conservation. Section~\ref{sec:level3}~~~ presents our numerical approach, detailing the simulation setup, analyzing the time evolution of electron velocity and kinetic energy, investigating the resonant conditions with wave angular momentum, and examining the energy transfer ratio and polarization characteristics. Section~\ref{sec:Discussion}~~~ provides a comprehensive discussion of the key findings and their physical implications. Finally, Section~\ref{sec:Conclusion}~~~ summarizes the principal conclusions and outlines potential directions for future research.

@@

Let's consider the process in which an electron cyclotron system under a uniform magnetic field emits a photon along z axis, as shown in Fig.~\ref{fig:Schematic}. The moving electron has the velocity $v_z$ along the background magnetic field and the $v_\perp$ cyclotron velocity. The kinetic energy along $z$ is $T = \gamma m_e c^2 - m_e c^2$, where $\gamma$ refers to the Lorentz factor. The internal energy represents as $U = \frac{1}{2} \gamma m_e v_\perp^2$.

@@

The $z$-component of the orbital angular momentum operator can be expressed in spherical coordinates as:
\begin{equation}
\hat{L}_z = -i\hbar\frac{\partial}{\partial\phi}  \label{eq:Lz}
\end{equation}
Combining Eq.~(\ref{eq:psi}) with Eq.~(\ref{eq:Lz}), we have 
\begin{equation}
-i\hbar \frac{\partial}{\partial \phi} \Psi = (p_\phi - eA_\phi) r \Psi
\end{equation}
As a result, the eigenvalue of  $L_z $can be expressed as
\begin{equation}
L_z = (p_\phi - e A_\phi) r  \label{eq:Lz_2}
\end{equation}
With \( p_\phi = \gamma m_e v_\perp \), \( A_\phi = \dfrac{r B_0}{2} \), and \( r = \dfrac{\gamma m_0 v_\perp}{B_0 e} \), Eq.~(\ref{eq:Lz_2}) can be rewritten as:
\begin{equation}
L_z = \frac{1}{2} \cdot \frac{\gamma m_0 v_\perp^2}{\omega_{ce}} = \frac{U}{\omega_{ce}},
\end{equation}
where \(\omega_{ce} = \dfrac{eB}{m_0 \gamma} = \dfrac{\omega_0}{\gamma}\) and \(U = \dfrac{1}{2} \gamma m_0 v_\perp^2\).
Here, \( m_0 \) is the electron rest mass, \( \gamma \) is the Lorentz factor, and \( \omega_0 \) is the electron cyclotron frequency in the rest frame (\( here we choose \omega_0 > 0 \)). The conservation of angular momentum in the \( z \)-direction is expressed as
\[
L_{z2} + m\hbar = L_{z1}.
\]
The variation in the angular momentum of the electron along the \( z \)-axis is given by:
\begin{equation}
\Delta L_{21} = L_{z2} - L_{z1} = \frac{U_2 - U_1}{\omega_{ce}} = -m\hbar \label{eq:DeltaL21}
\end{equation}
Here, \( m \) is the quantum number of the photon's angular momentum in the \( z \)-direction. The internal energy change is given by \( \Delta U_{21} = U_2 - U_1 \). With Eq.~(\ref{eq:DeltaL21}), it can be transformed as:
\begin{equation}
\Delta U_{21}=-m \hbar \omega_{ce}  \label{eq:DeltaU21}
\end{equation}
According to the Eq.~(\ref{eq:DeltaU00})  and Eq.~(\ref{eq:DeltaU21}), the change in electron energy  could be presented as 
\begin{equation}
\hbar \vec{k} \cdot \vec{v} = \hbar \omega - m \hbar \omega_{ce} \label{eq:EnergyResonant}
\end{equation}
This result is consistent with previous findings~\cite{tamm1959general, frank1960optics, nezlin1976negative, coppi1976slide, frolov1986excitation, ginzburg1996radiation}. Here, \( \hbar \vec{k} \cdot \vec{v} \) represents the loss of kinetic energy \( \Delta T_{12} \), \( \hbar \omega \) represents the energy of the photon, and \( -m \hbar \omega_{ce} \) represents the change in the electron gyrokinetic energy \( \Delta U_{21} \) (the internal energy change). The ratio between the internal energy change \( \Delta U_{21} \) and the kinetic energy change \( \Delta T_{21} \) can be expressed as
\begin{equation}
\frac{\Delta U_{21}}{\Delta T_{21}} = \frac{m \hbar \omega_{ce}}{\hbar \vec{k} \cdot \vec{v}}  \label{eq:DU21DT21}
\end{equation}
 This results is a critical criterion to compare with the classical dynamic simulation in the section 2. It is also proved based on classical theory in the Appendix.
After simpifying the Eq.~(\ref{eq:EnergyResonant}), we finally have the classical wave-particle resonant condition 
\begin{equation}
\omega = k_z v_z + m \omega_{ce}
\end{equation}
The variable \( m \) represents the quantum number associated with the angular momentum of the photon. Since a photon possesses both orbital angular momentum (\( l\hbar\), where \(l = 0,\pm1,\pm2,\pm3,... \)) and intrinsic spin angular momentum (\( s\hbar \), where \( s = \pm 1 \))~\cite{arnaut2000orbital}, the total angular momentum can be expressed as \( m\hbar = l\hbar + s\hbar \).

For photons carrying only spin angular momentum, two distinct quantum states are possible, characterized by the spin quantum number $m$:

\begin{enumerate}
\item For $m=+1$ ($\Delta U_{21}<0$), the cyclotron electron loses internal energy upon photon emission. The emitted photon displays right-hand circular polarization . This process is known as the NDE.

\item For $m=-1$ ($\Delta U_{21}>0$), the cyclotron electron gains internal energy through photon emission. The emitted photon exhibits left-hand circular polarization \footnote{The difference between our definition of circular polarization and the standard definition~\cite{kiang2008angular} stems from our choice of $\omega_0 > 0$. Here, $m > 0$ corresponds to the same rotational sense as the electron's natural right-hand gyration, which yields right-hand polarization when $\vec{k} \parallel \vec{B}_0$.}. This process corresponds to the ADE.
\end{enumerate}

%above describe spontaneous emission phenomenon of ADE and NDE without external field intervention . In our simulation model , an external plane EM field is introduced as resonant field interacting with electrons . the plane EM wave actually play a role as induced wave, where gyro-electron may absorb or stimulated em wave. from quantum from, the plane em wave actually is the ocean of photon which only contains spin angular momentum. so we should get the same resonant condition as derived from quantum field , where the righ hand polarization wave corresponds to $m = 1$ \cite{wei2024spin} and could only resonant under condition $\omega=\omega_c{e}+\vec{k}\cdot\vec{v}$,which is corresponds to NDR, while left-hand polarization corresponds to $m=-1$ and could only resonant with electron under $\omega=-\omega_{ce}+\vec{k}\cdot\vec{v}$,which is correspond to ADR
The above discussion described spontaneous emission phenomena of ADE and NDE without external field intervention. In our simulation model, we introduce an external plane electromagnetic (EM) wave that serves as a resonant field interacting with electrons. This plane EM wave acts as an inducing field, enabling gyro-electron to undergo stimulated absorption or emission processes. 
From a quantum perspective,  the plane EM wave can be regarded as an ocean of photons that carry only spin angular momentum. Consequently, the same resonance conditions as those derived from quantum field theory are recovered:

\begin{itemize}
    \item Right-hand circularly polarized  waves correspond to $m = +1$ states \cite{wei2024spin}, resonating only when $\omega = \omega_{ce} + \vec{k}\cdot\vec{v}$ (NDR condition)

    \item Left-hand circularly polarized  waves correspond to $m = -1$ states, resonating only when $\omega = -\omega_{ce} + \vec{k}\cdot\vec{v}$ (ADR condition)
\end{itemize}

This exact correspondence between our classical simulation framework and quantum field theoretic predictions validates our modeling approach while providing physical insight into the angular momentum selection rules governing these resonant interactions.

Although nonlinear analyses of electron interactions with electromagnetic waves have been extensively studied \cite{liu2004particle,qian1999exact,weyssow1990motion,gogoberidze2005origin,roberts1964motion,bourdier2000dynamics,nusinovich1999theory,nusinovich1995theory,qian2000relativistic}, the specific role of static electric fields in these interactions has received comparatively less attention. In our approach, the uniform electric field serves a crucial function by systematically scanning the electron velocity, thereby enabling investigation across the full spectrum of resonance conditions. The inherent complexity of these nonlinear processes precludes analytical solutions, necessitating the use of numerical simulation methods to obtain meaningful physical insights.

\section{Classical dynamic simulation of ADR}\label{sec:level3}
The ADE process has been analyzed based on quantum theory, demonstrating that the angular momentum of the emitting photon determines the resonance condition.  These characteristics will be tested through the interaction of EM wave and the electron during ADR and NDR, and the energy transfer ratio can also be verified through numerical simulations.
\subsection{Numerical simulation setup}
To analyze the resonant process from the perspective of classical dynamics and to provide a direct comparison between quantum and classical dynamic results, the following scenario is considered: A uniform magnetic field \( \vec{B}_0 \) is applied along the \( z \)-direction. An electrostatic field \( \vec{E}_0 \), oriented in the opposite direction to \( \vec{B}_0 \) (as illustrated in Fig.~\ref{fig:Setup}), is used to accelerate the electron. 
\begin{figure}
\centering
\includegraphics[width=0.4\textwidth]{Figure4.eps}% Here is how to import EPS art
\caption{\label{fig:Setup}The uniform static magnetic field is set along the z axis, the electrostatic field $E_0$ is oriented opposite to the $B_0$ field, and the wavevector $k$ is aligned parallel to the $B_0$ field.}
\end{figure}
We consider the interaction between an electron entering the system with velocity \( v_z \), parallel to the magnetic field \( B_0 = B_z \), and a linearly or circularly polarized transverse electromagnetic (TEM) wave propagating in a homogeneous dielectric medium with a refractive index \( n > 1 \).

The induced linearly polarized wave along \( \vec{B}_0 \) can be decomposed into a combination of a right-hand circularly polarized wave (\( m = 1 \)) and a left-hand circularly polarized wave (\( m = -1 \)), such that
\(
\vec{E}_w = \vec{E}_R + \vec{E}_L,
\)
where
\(
\vec{E}_R = \frac{1}{2} E_0 (\vec{e}_x + i\vec{e}_y) \exp[i(\vec{k} \cdot \vec{r} - \omega t)],
\vec{E}_L = \frac{1}{2} E_0 (\vec{e}_x - i\vec{e}_y) \exp[i(\vec{k} \cdot \vec{r} - \omega t)].
\)
%If the wavevector \( \vec{k} \) lies in the \( y \)-\( z \) plane with a crossing angle \( \theta_k \) relative to the \( z \)-axis, then the new coordinate unit vectors for the wave field expression should be rotated accordingly to align with the direction of \( \vec{k} \). The transformed basis vectors are:
%\begin{equation}
%\begin{pmatrix}
%\vec{e}_x' \\
%\vec{e}_y'
%\end{pmatrix}
%=
%\begin{pmatrix}
%1 & 0 & 0 \\
%0 & \cos\theta_k & \sin\theta_k
%\end{pmatrix}
%\cdot
%\begin{pmatrix}
%\vec{e}_x \\
%\vec{e}_y \\
%\vec{e}_z
%\end{pmatrix}
%\end{equation}
The magnetic field of E.M wave is
\begin{equation}
\vec{B}_w = \frac{\vec{k} \times \vec{E}_w}{\omega}
\end{equation}

@@

To simulate the evolution of $\bm{r}$ and $\bm{p}$, the above system is discretized using the Volume-Preserving Algorithm. Let $j$ denote the iteration step and $\text{Cay}(\bm{A})$ represent the Cayley transform of matrix $\bm{A}$:

\begin{equation}
\left\{
\begin{aligned}
\bm{r}^*_{j+\frac{1}{2}} &= \bm{r}^*_j + \frac{\Delta t^*}{2 \gamma_j} \bm{p}^*_j, \\
\bm{p}^{*-} &= \bm{p}^*_j + \frac{\Delta t^*}{2} \bm{E}^*_{j+\frac{1}{2}}, \\
\bm{p}^{*+} &= \text{Cay} \left( \frac{\Delta t^* \hat{\bm{B}}^*}{2 \gamma^{*-}} \right) \bm{p}^{*-}, \\
\bm{p}^*_{j+1} &= \bm{p}^{*+} + \frac{\Delta t^*}{2} \bm{E}^*_{j+\frac{1}{2}}, \\
\bm{r}^*_{j+1} &= \bm{r}^*_{j+\frac{1}{2}} + \frac{\Delta t^*}{2 \gamma_{j+1}} \bm{p}^*_{j+1},
\end{aligned}
\right.
\end{equation}
The dimensionless parameters are momentum \( p^* = p/(m_0 c) \), magnetic field \( B^* = B/(e \tau_{ce} m_0) \), total electric field \( E^* = E/(\frac{m_0 c}{\tau_{ce} e}) \), time step \( \Delta t^* = \Delta t / \tau_{ce} \), and position \( r^* = r / (\tau_{ce} c) \) respectively, where the \( \tau_{ce} \) is the electron cyclotron period (\( \tau_{ce} = 2\pi / \omega_{ce} \)) and \( \gamma^* = \sqrt{1 + p^{*2}} \) is Lorentz factor. The dimensionless magnetic matrix \( \mathbf{B}^* \) \cite{zhang2015volume} is written as

@@

The evolution of the electron’s motion is shown in Fig.~\ref{fig:kinetic_evolution}. As the electron accelerates from stationary in the electrostatic field (Fig.~\ref{fig:kinetic_evolution}(b)), the resonant frequencies increase simultaneously (Fig.~\ref{fig:kinetic_evolution}(a)). The change of parallel velocity caused by electromagnetic wave can be quantified as \( \Delta v = v_z - v_{zE0} \) as shown in Fig.~\ref{fig:kinetic_evolution}(c), where \( v_z \) represents the parallel velocity under the given scenario, while \( v_{zE0} \) denotes the parallel velocity resulting solely from the electrostatic field, which can be calculated using a theoretical equation as

@@

The cyclotron velocity is shown in Fig.~\ref{fig:kinetic_evolution}(d). The work done by electromagnetic wave is shown in fig. 5(e), which can be calculated by integrating the power with time as 
\(
E_{\parallel \text{emw}} = \int P_{\parallel \text{emw}} dt ,
\)
and 
\(
P_{\parallel \text{emw}} = -e \vec{v}_\perp \times \vec{B}_{\perp \text{emw}} \cdot v_z.
\)
Since all discrete data points are available from the simulation, it is not difficult to integrate all the discrete data over time. Fig.~\ref{fig:kinetic_evolution}(f) shows the gyro-kinetic energy evolution with time, where 
\(
E_\perp = \frac{1}{2} m_e v_\perp^2.
\)

\begin{figure}[ht]
\centering
\includegraphics[width=0.8\textwidth]{Figure5.pdf}% Here is how to import EPS art
\caption{\label{fig:NumSim}Kinetic evolution of electrons in a magnetic field with an electromagnetic wave during acceleration. 
    (a) Frequencies of ADE, NDE, and source wave frequency. 
    (b) The parallel velocity $v_z$ in the case with and without the electromagnetic wave. 
    (c) The change of parallel velocity caused by the electromagnetic wave. 
    (d) The cyclotron velocity $v_\perp$. 
    (e) The parallel kinetic energy transferred to the electron by the electromagnetic wave. 
    (f) The evolution of gyro-kinetic energy.}
    \label{fig:kinetic_evolution}
\end{figure}

@@

As shown in Fig.~\ref{fig:kinetic_evolution}(a), around \( 23\tau_{ce} \), the normal doppler frequency matches that of the induced wave, leading to a rapid increase in the cyclotron velocity \( v_{\perp} \) (Fig.~5(b)). Simultaneously, the change in parallel velocity induced by the electromagnetic wave also increases. This phenomenon can be interpreted as the electron cyclotron system absorbing a photon during the NDR, resulting in an increase in both parallel kinetic energy and gyro-kinetic energy(internal energy). The change in parallel kinetic energy caused by the electromagnetic wave is shown in Fig.~\ref{fig:kinetic_evolution}(e), where \( \Delta T_{21} = \Delta E_{||\text{emw}} \approx 5.4 \times 10^{-5} \, \text{eV} \). The increase in gyro-kinetic energy is shown in Fig.~\ref{fig:kinetic_evolution}(e), where \( \Delta U_{21} = \Delta E_{\perp} \approx 10.7 \times 10^{-5} \, \text{eV} \). The energy transfer ratio between internal energy and parallel kinetic energy during resonance is given by \( \frac{\Delta U_{21}}{\Delta T_{21}} \approx 1.98 \). According to quantum theory, the energy ratio is given by Eq.~(13). Here \( m = 1 \) for NDE and \( k = 10^5 \, \text{m}^{-1} \) along the \( z \)-axis, the resonant velocity \( v_z \approx 19 \times 10^3 \, \text{m/s} \) and \( \omega_{ce} \approx 3.51 \times 10^9 \, \text{s}^{-1} \). Finally, \( n_p = 1.85 \), which is in close agreement with the simulation results.
